---
title: "DATA 624 Assignment 2"
author: "Philip Tanofsky"
date: "9/19/2021"
output: html_document
---

```{r warning=F, message=F}
# Import required R libraries
library(fpp3)
```
 
# Exercise 3.1

Consider the GDP information in `global_economy`. Plot the GDP per capita for each country over time. Which country has the highest GDP per capita? How has this changed over time?

```{r warning=F, message=F}

global_economy %>%
  autoplot(GDP/Population, show.legend=FALSE) +
  labs(title= "GDP per capita", y = "$US")
```

# Exercise 3.2

For each of the following series, make a graph of the data. If transforming seems appropriate, do so and describe the effect.

## A

United States GDP from `global_economy`

```{r warning=F, message=F}

global_economy %>%
  filter(Country == "United States") %>%
  autoplot(GDP/Population) +
  labs(title= "GDP per capita", y = "$US")
```

Population adjustment

## B

Slaughter of Victorian "Bulls, bullocks and steers" in `aus_livestock`.

```{r warning=F, message=F}

aus_livestock %>%
  filter(Animal == "Bulls, bullocks and steers" &
           State == "Victoria") %>%
  autoplot(Count) +
  labs(title= "Bulls, bullocks and steers", y = "Count")
```

## C

Victorian Electricity Demand from `vic_elec`.

```{r warning=F, message=F}

vic_elec %>%
  autoplot(Demand) +
  labs(title= "Electricity Demand", y = "MW")
```

## D

Gas production from `aus_production`.

```{r warning=F, message=F}

aus_production %>%
  autoplot(Gas) +
  labs(title= "Gas Production", y = "Petajoule")
```
Box-Cox (mathematical transformation)


# Exercise 3.3

Why is a Box-Cox transformation unhelpful for the `canadian_gas` data?

```{r warning=F, message=F}

canadian_gas %>%
  autoplot(Volume) +
  labs(title= "Monthly Canadian Gas Production", y = "Billions of cubic meters")
```

# Exercise 3.4

What Box-Cox transformation would you select for your retail data (from Exercise 8 in Section 2.10)?

```{r warning=F, message=F}
set.seed(8675309)

myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))

head(myseries)

dim(myseries)
```

# Exercise 3.5

For the following series, find an appropriate Box-Cox transformation in order to stabilize the variance. Tobacco from `aus_production`, Economy class passengers between Melbourne and Sydney from `ansett`, and Pedestrian counts at Southern Cross Station from `pedestrian`.

## A - aus_production

```{r warning=F, message=F}

```

## B - ansett

```{r warning=F, message=F}

```

## C - pedestrian

```{r warning=F, message=F}

```

# Exercise 3.7

Consider the last five years of the Gas data from `aus_production`.

```{r warning=F, message=F}
gas <- tail(aus_production, 5*4) %>% select(Gas)
```

## A

Plot the time series. Can you identify seasonal fluctuations and/or a trend-cycle?

## B

Use `classical_decomposition` with `type=multiplicative` to calculate the trend-cycle and seasonal indices.

## C

Do the results support the graphical interpretation from part a?

## D

Compute and plot the seasonally adjusted data.

## E

Change one observation to be an outlier (e.g., add 300 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

## F

Does it make any difference if the outlier is near the end rather than in the middle of the time series?

# Exercise 3.8

Recall your retail time series data (from Exercise 8 in Section 2.10). Decompose the series using X-11. Does it reveal any outliers, or unusual features that you had not noticed previously?

# Exercise 3.9

Figures 3.19 and 3.20 show the result of decomposing the number of persons in the civilian labour force in Australia each month from February 1978 to August 1995.

## A

Write about 3–5 sentences describing the results of the decomposition. Pay particular attention to the scales of the graphs in making your interpretation.

## B

Is the recession of 1991/1992 visible in the estimated components?






```{r warning=F, message=F}

```

