---
title: "DATA 624 Assignment 10"
subtitle: "CUNY Fall 2021"
author: "Philip Tanofsky"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: false
    theme: united
    highlight: tango
---

# Prompt

Imagine 10000 receipts sitting on your table. Each receipt represents a transaction with items that were purchased. The receipt is a representation of stuff that went into a customer's basket - and therefore 'Market Basket Analysis'.

That is exactly what the Groceries Data Set contains: a collection of receipts with each line representing 1 receipt and the items purchased. Each line is called a transaction and each column in a row represents an item. The data set is attached.

Your assignment is to use R to mine the data for association rules. You should report support, confidence and lift and your top 10 rules by lift.

**Extra credit:** do a simple cluster analysis on the data as well. Use whichever packages you like. Due May 3 before midnight.

```{r warning=F, message=F}
# Import required R libraries
#library(tidyverse)
library(arules)
library(arulesViz)

# Set seed for assignment
set.seed(200)
```

The `arules` package (<https://www.rdocumentation.org/packages/arules/versions/1.7-1>) "provides the infrastructure for representing, manipulating and analyzing transaction data and patterns using frequent itemsets and association rules." The `arules` library contains the data structure definitions and mining algorithms - APRIORI and ECLAT.

The `arulesViz` library provides visualizations for the association rules.

# Exploratory Data Analysis

Using the example provided at <http://r-statistics.co/Association-Mining-With-R.html>, I read in the provided CSV file as transactions objects per the `arules` package.

```{r warning=F, message=F}
# http://r-statistics.co/Association-Mining-With-R.html
grocery_ds <- read.transactions("GroceryDataSet.csv", sep=",")
class(grocery_ds)
```

The `class()` function confirms the `grocery_ds` object is transactions from the `arules` package.

```{r warning=F, message=F}
summary(grocery_ds)
```

Summary of the grocery data set indicates 9835 rows, or individual receipts defined as transactions, with a total of 169 columns, or unique items. The most frequent items are whole milk, other vegetables, rolls/buns, soda, and yogurt. The median number of items per transaction is 3 with an average number of 4.4 items per transaction. The minimum is 1, which is expected, and the maximum items per a transaction is 32.

```{r warning=F, message=F}
size(head(grocery_ds))
LIST(head(grocery_ds, 3))
inspect(head(grocery_ds, 3))
```

The `size()` function confirms the item count per transaction and the `LIST()` function confirms the data object meets expectations, the first three transactions are displayed. The results above match the raw CSV file. The `inspect()` functions appears to behave the same as `LIST()` but perhaps a cleaner output presentation.

## Support Evaluation

```{r warning=F, message=F}
# calculates 'support' of the frequent items in the dataset
support_val <- 0.07
frequentItems <- eclat(grocery_ds, parameter = list(supp=support_val, maxlen=15))
```

The `eclat()` function "finds frequent item sets with the Eclat algorithm, which carries out a depth first search on the subset lattice and determines the support of item sets by intersecting transaction lists." (<https://borgelt.net/eclat.html>) The parameter `supp` represents $support$ which is defined as the proportion of transactions in the dataset which contain the item. The parameter value is the minimum threshold.

```{r warning=F, message=F}
inspect(frequentItems)
```

With a support value of `r support_val`, the output reports the above 19 results which indicates only one pair of items appear at the given proportion of the time, whole milk and other vegetables. The result makes sense given those two items are the most frequent individually and are grocery staples.

```{r warning=F, message=F}
itemFrequencyPlot(grocery_ds, topN=10, type="absolute", main="Item Frequency")
```

The plot above displays a count of the 10 most frequent items with whole milk and other vegetables occurring most often, matching the results of the `summary()` function above.

# Generate Association Rules

## Confidence Evaluation

```{r warning=F, message=F}
# Define minimum support
supp_val <- 0.001
# Define minimum confidence (increase to get stronger rules)
conf_val <- 0.9
# Increase maxlen to get longer rules
maxlen_val <- 5
rules <- apriori(grocery_ds, parameter=list(supp=supp_val, conf=conf_val, maxlen=maxlen_val))
rules_conf <- sort(rules, by="confidence", decreasing=TRUE)
```

The `apriori()` function "finds association rules and frequent item sets with the Apriori algorithm, which carries out a breadth first search on the subset lattice and determines the support of item sets by subset tests." (<https://borgelt.net/apriori.html>)

```{r warning=F, message=F}
inspect(head(rules_conf, 10))
```

With a support value of `r supp_val` and confidence value of `r conf_val`, the above output shows 10 association rules with a resulting confidence value of 1. The confidence value of 1 indicates the item on the right hand side always occurs when the item or items on the left hand side occur. Not surprisingly, of the 10 rules displayed above, 9 rules indicate whole milk on the right hand side with other vegetables as the remaining value on the right hand side. Given the higher frequency of those two items, these results are expected.

## Lift Evaluation

```{r warning=F, message=F}
rules_lift <- sort(rules, by="lift", decreasing=TRUE)
inspect(head(rules_lift, 10))
```

The same values of support and confidence, and thus from the same resulting rules, the rules are sorted by lift value in order to find the 10 rules with the highest lift. The lift value indicates "the deviation of the support of the whole rule from the support expected under independence given the supports of the LHS and the RHS." (https://cran.r-project.org/web/packages/arules/vignettes/arules.pdf) The higher the lift value, the stronger the association between the LHS and the RHS.

Note: I decided on confidence value of `r conf_val` so I could receive some high lift rules that didn't indicate whole milk and other vegetables on the right hand side. The top lift result (11.235269) given the parameter values is an LHS of liquor and red/blush wine and an RHS of bottled beer. That association makes sense for someone buying alcohol together. The rules with an RHS of yogurt shows different LHS items including butter, cream cheese, curd, and whipped/sour cream which seems valid given those items are typically in the same refrigerated section of a grocery store. Overall, none of the above rules sticks out like the infamous diapers to beer association.

```{r warning=F, message=F}
# Could I find the diapers => beer rule
supp_baby_val <- 0.001
conf_baby_val <- 0.05
rules_baby <- apriori(grocery_ds, 
                 parameter=list(supp=supp_baby_val, conf=conf_baby_val),
                 appearance = list (default="rhs", lhs="baby food"),
                 control = list (verbose=F))

rules_baby_conf <- sort (rules_baby, by="confidence", decreasing=TRUE)
inspect(head(rules_baby_conf))
```

Nope, couldn't find the baby product to beer rule. Apparently, "baby food" as the LHS didn't produce any meaningful association, nor did "baby cosmetics" as the LHS.

# Visualizations

I used the plotting functions from the library `arulesViz` to help understand the association rules through visualizations.

```{r warning=F, message=F}
# https://cran.r-project.org/web/packages/arulesViz/vignettes/arulesViz.pdf
options(digits = 2)
plot(rules)
```

The above scatterplot indicates the relationship the values of Support and Confidence for the 123 association rules generated by with a support value of `r supp_val` and confidence value of `r conf_val`. 

```{r warning=F, message=F}
plot(rules, measure=c("support", "lift"), shading="confidence")
```

The above scatterplot indicates the relationship the values of Support and Lift for the 123 association rules generated by with a support value of `r supp_val` and confidence value of `r conf_val`. 

```{r warning=F, message=F}
plot(rules, method="two-key plot")
```

The above two-key scatterplot indicates the relationship the values of Support and Confidence for the 123 association rules generated by with a support value of `r supp_val` and confidence value of `r conf_val` in which order identifies the number of items in the rule.

```{r warning=F, message=F, fig.height=8}
plot(rules, method="grouped", control = list(k = 10))
```

The above grouped matrix-based visualization uses a balloon plot to show the LHS values as columns and the RHS items as rows. The color of the balloon shows the aggregated interest measure and the size of the balloon show the aggregated support.

```{r warning=F, message=F}
subrules2 <- head(rules, n=10, by="lift")
plot(subrules2, method="graph")
```

The above graph-based visualization shows the items and rules as vertices and connections with directed edges. The plot helps identify which rules share items.

# Conclusion

Overall, the market basket analysis proved straightforward with the use of the `arules` package. In order to tease out some "interesting" associations, then more modifications of the support and confidence levels would be required to find associations with few occurrences but with high confidence values. Too bad this dataset didn't have the diapers to beer connection.

