---
title: "DATA 624 Assignment 6"
subtitle: "CUNY Fall 2021"
author: "Philip Tanofsky"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=F, message=F}
# Import required R libraries
library(fpp3)
```


# Exercise 9.1

Figure 9.32 shows the ACFs for 36 random numbers, 360 random numbers and 1,000 random numbers.

### Section a

Explain the differences among these figures. Do they all indicate that the data are white noise?

**Answer:** XXX

### Section b

Why are the critical values at different distances from the mean of zero? Why are the autocorrelations different in each figure when they each refer to white noise?

**Answer:** XXX

# Exercise 9.2

A classic example of a non-stationary series are stock prices. Plot the daily closing prices for Amazon stock (contained in `gafa_stock`), along with the ACF and PACF. Explain how each plot shows that the series is non-stationary and should be differenced.

```{r warning=F, message=F}
gafa_stock %>%
  filter(Symbol == 'AMZN') %>%
  autoplot(Close) +
  labs(title = "Amazon Daily Closing Stock Price",
       y = "Price")
```

```{r warning=F, message=F}
gafa_stock %>%
  filter(Symbol == "AMZN") %>%
  ACF(Close) %>%
  autoplot()
```

```{r warning=F, message=F}
gafa_stock %>%
  filter(Symbol == "AMZN") %>%
  PACF(Close) %>%
  autoplot()
```

# Exercise 9.3

For the following series, find an appropriate Box-Cox transformation and order of differencing in order to obtain stationary data.

### Section a

Turkish GDP from `global_economy`.

```{r warning=F, message=F}
turkey_gdp <- global_economy %>%
  filter(Country == 'Turkey')

lambda <- turkey_gdp %>%
  features(GDP, features = guerrero) %>%
  pull(lambda_guerrero)

turkey_gdp %>%
  autoplot(box_cox(GDP, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Transformed Turkish GDP with $\\lambda$ = ",
         round(lambda,2))))
```

Appears to have an increasing trend after Box-Cox transformation. Attempt first differencing.

```{r warning=F, message=F}
turkey_gdp <- turkey_gdp %>%
  mutate(diff_bc_gdp = difference(box_cox(GDP, lambda)))

# Display plot of first difference
turkey_gdp %>%
  autoplot(diff_bc_gdp) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Transformed Turkish GDP with $\\lambda$ = ",
         round(lambda,2))))
```

```{r warning=F, message=F}
# Apply Ljung-Box test
turkey_gdp %>%
  features(diff_bc_gdp, ljung_box, lag = 10)
```

### Section b

Accommodation takings in the state of Tasmania from `aus_accommodation`.

Quarterly data

```{r warning=F, message=F}
tas_takings <- aus_accommodation %>%
  filter(State == 'Tasmania')

lambda <- tas_takings %>%
  features(Takings, features = guerrero) %>%
  pull(lambda_guerrero)
```
tas_takings %>%
  autoplot(box_cox(Takings, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Transformed gas production with $\\lambda$ = ",
         round(lambda,2))))

```{r warning=F, message=F}
tas_takings %>%
  transmute(
    `Takings` = Takings,
    `Box-Cox Takings` = box_cox(Takings, lambda),
    `Annual change in Box-Cox Takings` = difference(box_cox(Takings, lambda), 4),
    `Doubly differenced Takings` =
                     difference(difference(box_cox(Takings, lambda), 4), 1)
  ) %>%
  pivot_longer(-Date, names_to="Type", values_to="Takings") %>%
  mutate(
    Type = factor(Type, levels = c(
      "Takings",
      "Box-Cox Takings",
      "Annual change in Box-Cox Takings",
      "Doubly differenced Takings"))
  ) %>%
  ggplot(aes(x = Date, y = Takings)) +
  geom_line() +
  facet_grid(vars(Type), scales = "free_y") +
  labs(title = "Tasmanian Accomodation Takings", y = NULL)
```

Definitely appears to have a seasonal component.

Appears the doubly differencing is necessary.

```{r warning=F, message=F}
# Apply Ljung-Box test
tas_takings %>%
  features(difference(difference(box_cox(Takings, lambda), 4), 1), ljung_box, lag = 12)

## NOT THE RESULT I WANT, DOUBLE-CHECK ##

tas_takings %>%
  ACF(difference(difference(box_cox(Takings, lambda), 4), 1)) %>%
  autoplot()
```

### Section c

Monthly sales from `souvenirs`.

```{r warning=F, message=F}
lambda <- souvenirs %>%
  features(Sales, features = guerrero) %>%
  pull(lambda_guerrero)

souvenirs %>%
  autoplot(box_cox(Sales, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Transformed gas production with $\\lambda$ = ",
         round(lambda,2))))

souvenirs %>%
  transmute(
    `Sales` = Sales,
    `Box-Cox sales` = box_cox(Sales, lambda),
    `Annual change in Box-Cox sales` = difference(box_cox(Sales, lambda), 12),
    `Doubly differenced Box-Cox sales` =
                     difference(difference(box_cox(Sales, lambda), 12), 1)
  ) %>%
  pivot_longer(-Month, names_to="Type", values_to="Sales") %>%
  mutate(
    Type = factor(Type, levels = c(
      "Sales",
      "Box-Cox sales",
      "Annual change in Box-Cox sales",
      "Doubly differenced Box-Cox sales"))
  ) %>%
  ggplot(aes(x = Month, y = Sales)) +
  geom_line() +
  facet_grid(vars(Type), scales = "free_y") +
  labs(title = "Souvenirs Sales", y = NULL)
```

Definitely appears to have a seasonal component.

Appears the double differencing is needed.

```{r warning=F, message=F}
# Apply Ljung-Box test
souvenirs %>%
  features(difference(difference(box_cox(Sales, lambda), 12), 1), ljung_box, lag = 12)

## NOT THE RESULT I WANT, DOUBLE-CHECK ##
```

# Exercise 9.5

For your retail data (from Exercise 8 in Section 2.10), find the appropriate order of differencing (after transformation if necessary) to obtain stationary data.

```{r warning=F, message=F}
set.seed(8675309)

myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))
```

# Exercise 9.6

Simulate and plot some data from simple ARIMA models.

### Section a

Use the following R code to generate data from an AR(1) model with  
$\phi_1 = 0.6$ and $\sigma^2 = 1$. The process starts with $y_1 = 0$.
 
```{r warning=F, message=F}
y <- numeric(100)
e <- rnorm(100)
for(i in 2:100)
  y[i] <- 0.6*y[i-1] + e[i]
sim <- tsibble(idx = seq_len(100), y = y, index = idx)
```

### Section b

Produce a time plot for the series. How does the plot change as you change $\phi_1$?

```{r warning=F, message=F}

```

### Section c

Write your own code to generate data from an MA(1) model with $\theta_1 = 0.6$ and $\sigma^2 = 1$.

```{r warning=F, message=F}

```

### Section d

Produce a time plot for the series. How does the plot change as you change $\theta_1$?

```{r warning=F, message=F}

```

### Section e

Generate data from an ARMA(1,1) model with  $\phi_1 = 0.6,$ $\theta_1 = 0.6$ and $\sigma^2 = 1$.

```{r warning=F, message=F}

```

### Section f

Generate data from an AR(2) model with $\phi_1 = -0.8$, $\phi_2 = 0.3$ and $\sigma^2 = 1$. (Note that these parameters will give a non-stationary series.)

```{r warning=F, message=F}

```

### Section g

Graph the latter two series and compare them.

```{r warning=F, message=F}

```

# Exercise 9.7

Consider `aus_airpassengers`, the total number of passengers (in millions) from Australian air carriers for the period 1970-2011.

```{r warning=F, message=F}
aus_airpassengers %>% head()
```

### Section a

Use `ARIMA()` to find an appropriate ARIMA model. What model was selected. Check that the residuals look like white noise. Plot forecasts for the next 10 periods.

```{r warning=F, message=F}
# Generate ARIMA model
aus_air_mod <- aus_airpassengers %>% model(ARIMA(Passengers))
```

Display model definition

```{r warning=F, message=F}
# Output report to show model selected
report(aus_air_mod)
```

Check the residuals look like white noise

```{r warning=F, message=F}
aus_air_mod %>% gg_tsresiduals()
```

Plot forecasts for the next 10 periods

```{r warning=F, message=F}
aus_air_fc <- aus_air_mod %>% forecast(h=10)

aus_air_fc
```

### Section b

Write the model in terms of the backshift operator.

```{r warning=F, message=F}

```

### Section c

Plot forecasts from an ARIMA(0,1,0) model with drift and compare these to part a.

```{r warning=F, message=F}

```

### Section d

Plot forecasts from an ARIMA(2,1,2) model with drift and compare these to parts a and c. Remove the constant and see what happens.

```{r warning=F, message=F}

```

### Section e

Plot forecasts from an ARIMA(0,2,1) model with a constant. What happens?

```{r warning=F, message=F}

```

# Exercise 9.8 (confirm before working on)




```{r warning=F, message=F}

```

```{r warning=F, message=F}

```

```{r warning=F, message=F}

```











