---
title: "DATA 624 Assignment 6"
subtitle: "CUNY Fall 2021"
author: "Philip Tanofsky"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=F, message=F}
# Import required R libraries
library(fpp3)
```


# Exercise 9.1

Figure 9.32 shows the ACFs for 36 random numbers, 360 random numbers and 1,000 random numbers.

### Section a

Explain the differences among these figures. Do they all indicate that the data are white noise?

**Answer:** XXX

### Section b

Why are the critical values at different distances from the mean of zero? Why are the autocorrelations different in each figure when they each refer to white noise?

**Answer:** XXX

# Exercise 9.2

A classic example of a non-stationary series are stock prices. Plot the daily closing prices for Amazon stock (contained in `gafa_stock`), along with the ACF and PACF. Explain how each plot shows that the series is non-stationary and should be differenced.

```{r warning=F, message=F}

```

# Exercise 9.3

For the following series, find an appropriate Box-Cox transformation and order of differencing in order to obtain stationary data.

### Section a

Turkish GDP from `global_economy`.

```{r warning=F, message=F}

```

### Section b

Accommodation takings in the state of Tasmania from `aus_accommodation`.

```{r warning=F, message=F}

```

### Section c

Monthly sales from `souvenirs`.

```{r warning=F, message=F}

```

# Exercise 9.5

For your retail data (from Exercise 8 in Section 2.10), find the appropriate order of differencing (after transformation if necessary) to obtain stationary data.

# Exercise 9.6

Simulate and plot some data from simple ARIMA models.

```{r warning=F, message=F}

```

### Section a

Use the following R code to generate data from an AR(1) model with  
$\phi_1 = 0.6$ and $\sigma^2 = 1$. The process starts with $y_1 = 0$.
 
```{r warning=F, message=F}
y <- numeric(100)
e <- rnorm(100)
for(i in 2:100)
  y[i] <- 0.6*y[i-1] + e[i]
sim <- tsibble(idx = seq_len(100), y = y, index = idx)
```

### Section b

Produce a time plot for the series. How does the plot change as you change $\phi_1$?

```{r warning=F, message=F}

```

### Section c

Write your own code to generate data from an MA(1) model with $\theta_1 = 0.6$ and $\sigma^2 = 1$.

```{r warning=F, message=F}

```

### Section d

Produce a time plot for the series. How does the plot change as you change $\theta_1$?

```{r warning=F, message=F}

```

### Section e

Generate data from an ARMA(1,1) model with  $\phi_1 = 0.6,$ $\theta_1 = 0.6$ and $\sigma^2 = 1$.

```{r warning=F, message=F}

```

### Section f

Generate data from an AR(2) model with $\phi_1 = -0.8$, $\phi_2 = 0.3$ and $\sigma^2 = 1$. (Note that these parameters will give a non-stationary series.)

```{r warning=F, message=F}

```

### Section g

Graph the latter two series and compare them.

```{r warning=F, message=F}

```

# Exercise 9.7

Consider `aus_airpassengers`, the total number of passengers (in millions) from Australian air carriers for the period 1970-2011.

### Section a

Use `ARIMA()` to find an appropriate ARIMA model. What model was selected. Check that the residuals look like white noise. Plot forecasts for the next 10 periods.

```{r warning=F, message=F}

```

### Section b

Write the model in terms of the backshift operator.

```{r warning=F, message=F}

```

### Section c

Plot forecasts from an ARIMA(0,1,0) model with drift and compare these to part a.

```{r warning=F, message=F}

```

### Section d

Plot forecasts from an ARIMA(2,1,2) model with drift and compare these to parts a and c. Remove the constant and see what happens.

```{r warning=F, message=F}

```

### Section e

Plot forecasts from an ARIMA(0,2,1) model with a constant. What happens?

```{r warning=F, message=F}

```

# Exercise 9.8 (confirm before working on)




```{r warning=F, message=F}

```

```{r warning=F, message=F}

```

```{r warning=F, message=F}

```











