---
title: "Algorithm Attempt v1"
author: "Philip Tanofsky"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: false
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, error=FALSE}
# Required libraries
library(tidyverse)
library(hoopR)
library(readxl)
library(writexl)
library(lubridate)
library(RMariaDB)
```

```{r}
DEBUG_PRINTS <- FALSE
```

# Introduction

```{r}
# FUNCTIONS

# ===== XXX =====
get_todays_date <- function() {
  today <- Sys.Date()
  today <- gsub("-", "", today)
  return(today)
}

# ===== XXX =====
get_yesterdays_date <- function() {
  yday <- Sys.Date()-1
  yday <- gsub("-", "", yday)
  return(yday)
}

convert_date_format <- function(date) {
  date_t <- date
  date_t <- gsub("-", "", date)
  return(date_t)
}

# ===== XXX =====
get_nba_games <- function(date) {
  games <- espn_nba_scoreboard (season = date)
}

# ===== XXX =====
convert_team_abbr <- function(name) {
  
  abbr <- "0"
  
  if (name == "NO") {
    abbr <- "NOP"
  } else if (name == "WSH") {
    abbr <- "WAS"
  } else if (name == "SA") {
    abbr <- "SAS"
  } else if (name == "UTAH") {
    abbr <- "UTA"
  } else if (name == "NY") {
    abbr <- "NYK"
  } else if (name == "GS") {
    abbr <- "GSW"
  } else {
    abbr <- name
  }
  
  return(abbr)
}

# ===== XXX =====
convert_results_format <- function(games, date) {
  games_l <- games
  games_l <- games %>% select(c(away_team_abb, away_score, home_team_abb, home_score, game_id))
  
  # Convert the abbreviations
  for(i in 1:nrow(games_l)) {
    games_l[i,]$away_team_abb <- convert_team_abbr(games_l[i,]$away_team_abb)
    games_l[i,]$home_team_abb <- convert_team_abbr(games_l[i,]$home_team_abb)
  }
  
  
  date_formatted <- ymd(as.numeric(date))
  games_l$DATE <- date_formatted
  games_l <- games_l %>% rename(c(DATE=DATE, 
                                  GAME_ID=game_id, 
                                  AWAY_TEAM=away_team_abb, 
                                  AWAY_TEAM_SCORE=away_score, 
                                  HOME_TEAM=home_team_abb, 
                                  HOME_TEAM_SCORE=home_score))
  games_l <- games_l[, c("DATE", "GAME_ID", "AWAY_TEAM", "AWAY_TEAM_SCORE", "HOME_TEAM", "HOME_TEAM_SCORE")]
  return(games_l)
}

# ===== XXX =====
add_game_results <- function(games) {
  games_from_file <- read_excel("NbaGames_v1.xlsx")
  (nrow(games_from_file))
  games_bind <- rbind(games_from_file, games)
  (nrow(games_bind))
  write_xlsx(games_bind, "NbaGames_v1.xlsx")
}

add_game_results_to_db <- function(games_df) {
  
  localuserpassword <- "BostonCeltics17"
  
  # The connection method below uses a password stored in a variable.
  # To use this set localuserpassword="The password of newspaper_search_results_user"
  nbaDb <- dbConnect(MariaDB(), 
                     user='nba_game_statistics_user', 
                     password=localuserpassword, 
                     dbname='nba_game_statistics', 
                     host='localhost')

  dbWriteTable(nbaDb, "schedule", games_df, append=TRUE ) 

  # Disconnect to clean up the connection to the database.
  dbDisconnect(nbaDb)
}

# ===== XXX =====
determine_game_winner <- function(home_score, away_score) {
  
  # Default to E for error
  result <- "E"
  
  if (home_score > away_score) {
    result <- "H"
  } else if (away_score > home_score) {
    result <- "A"
  }
  
  return(result)
}

# ===== XXX =====
calculate_e_team <- function(team_elo, opp_elo) {
  
  dividend <- 1 + 10^((opp_elo - team_elo) / 400)
  result <- 1 / dividend
  return(result)
}

# ===== XXX =====
calculate_k <- function(mov_winner, elo_diff_winner) {
  
  k <- 0
  dividend <- 0
  divisor <- 0
  
  dividend <- 7.5 + 0.006 * elo_diff_winner
  
  divisor <- (mov_winner + 3)^(0.8)
  
  k <- 20 * divisor / dividend
  
  return(k)
}

# ===== XXX =====
calculate_elo_iplus1 <- function(k, s_team, e_team, elo_i) {
  result <- k * (s_team - e_team) + elo_i
  return(result)
}

# ===== XXX =====
calculate_elo_ratings <- function(home_elo, away_elo, home_score, away_score) {
  # Add 100 to home elo to account for home court advantage
  HOME_COURT_ADV <- 100
  
  team_winner <- determine_game_winner(home_score, away_score)
  
  # Same no matter the winner
  e_team_home <- calculate_e_team(home_elo+HOME_COURT_ADV, away_elo)
  e_team_away <- calculate_e_team(away_elo, home_elo+HOME_COURT_ADV)
  
  k <- -1
  s_home <- -1
  s_away <- -1
  mov_multiplier <- -1
  
  if (team_winner == "H") {
    # Home team won game
    k <- calculate_k(home_score-away_score, home_elo+HOME_COURT_ADV-away_elo)
    s_home <- 1
    s_away <- 0
  } else if (team_winner == "A") {
    # Away team won game
    k <- calculate_k(away_score-home_score, away_elo-home_elo+HOME_COURT_ADV)
    s_home <- 0
    s_away <- 1
  } else {
    # Error has occured ... tie game?
    print("ERROR HAS OCCURRED: Tie Game")
  }
  
  if (DEBUG_PRINTS) {
    print(k)
    print(s_away)
    print(e_team_away)
    print(away_elo)
  }
  
  home_elo_iplus1 <- calculate_elo_iplus1(k, s_home, e_team_home, home_elo)
  away_elo_iplus1 <- calculate_elo_iplus1(k, s_away, e_team_away, away_elo)
  
  # Round to thousandths place
  home_elo_iplus1 <- trunc(home_elo_iplus1*10^4)/10^4
  away_elo_iplus1 <- trunc(away_elo_iplus1*10^4)/10^4
  
  return(list(home_elo_iplus1, away_elo_iplus1))
}
```

```{r}
# Validate above functions

# variables for testing
away_elo_i = 1597
home_elo_i = 1646
away_game_score = 116
home_game_score = 109

(determine_game_winner(home_game_score, away_game_score))
(determine_game_winner(away_game_score, home_game_score))

(calculate_elo_iplus1(16.165958095828678, 1, 0.29781739790947936, away_elo_i))

(calculate_e_team(home_elo_i+100, away_elo_i))
(calculate_e_team(away_elo_i, home_elo_i+100))

(calculate_k(away_game_score-home_game_score, away_elo_i-home_elo_i+100))
```


```{r}


```

```{r}
# Get today's data and remove dashes
date <- get_yesterdays_date()

date <- '20211019'

# Get NBA games for data (today in this case)
games <- get_nba_games(date)
#games

games_df <- convert_results_format(games, date)
games_df


localuserpassword <- "BostonCeltics17"
# The connection method below uses a password stored in a variable.
# To use this set localuserpassword="The password of newspaper_search_results_user"

nbaDb <- dbConnect(MariaDB(), user='nba_game_statistics_user', password=localuserpassword, dbname='nba_game_statistics', host='localhost')

dbWriteTable(nbaDb, "schedule", games_df, append=TRUE ) 

# Disconnect to clean up the connection to the database.

dbDisconnect(nbaDb)


# Add latest games to running file
#add_game_results(games_df)
```

#30#




```{r}
# Get scores from day before
# Add to Excel of all games
# Output total number of games/rows before and after
# DATE - AWAY_TEAM  - AWAY_SCORE - HOME_TEAM - HOME_SCORE - GAME_ID

# Update ELO rating

# Input odds
dbDisconnect(nbaDb)
```


```{r}
# GSW vs BKN
# Game played on December 22, 2020: Opening night
gsw_elo_i = 1395
bkn_elo_i = 1489
gsw_score = 99
bkn_score = 125

home1_away2 <- calculate_elo_ratings(bkn_elo_i, gsw_elo_i, bkn_score, gsw_score)

print("Home - BKN: ")
print(home1_away2[1])

print("Away - GSW: ")
print(home1_away2[2])
```


```{r}

# This got all the current results into the table
for(i in 1) {
  date_full <- Sys.Date()-i
  date <- convert_date_format(date_full)
  print(date)
  
  if (date != '20211125' && date != '20211224') {
    games <- get_nba_games(date)
    games_df <- convert_results_format(games, date)
    add_game_results_to_db(games_df)
  }
}
```


LOCK TABLES `elo_rating` WRITE;

# Added the initial score for each team
INSERT INTO nba_game_statistics.elo_rating VALUES
('YEARSTARTATL', '2021-10-19', 'ATL', '1570.000'),
('YEARSTARTBOS', '2021-10-19', 'BOS', '1500.000'),
('YEARSTARTBKN', '2021-10-19', 'BKN', '1605.000'),
('YEARSTARTCHA', '2021-10-19', 'CHA', '1438.000'),
('YEARSTARTCHI', '2021-10-19', 'CHI', '1495.000'),
('YEARSTARTCLE', '2021-10-19', 'CLE', '1353.000'),
('YEARSTARTDAL', '2021-10-19', 'DAL', '1541.000'),
('YEARSTARTDEN', '2021-10-19', 'DEN', '1577.000'),
('YEARSTARTDET', '2021-10-19', 'DET', '1381.000'),
('YEARSTARTGSW', '2021-10-19', 'GSW', '1529.000'),
('YEARSTARTHOU', '2021-10-19', 'HOU', '1346.000'),
('YEARSTARTIND', '2021-10-19', 'IND', '1487.000'),
('YEARSTARTLAC', '2021-10-19', 'LAC', '1624.000'),
('YEARSTARTLAL', '2021-10-19', 'LAL', '1549.000'),
('YEARSTARTMEM', '2021-10-19', 'MEM', '1541.000'),
('YEARSTARTMIA', '2021-10-19', 'MIA', '1513.000'),
('YEARSTARTMIL', '2021-10-19', 'MIL', '1658.000'),
('YEARSTARTMIN', '2021-10-19', 'MIN', '1439.000'),
('YEARSTARTNOP', '2021-10-19', 'NOP', '1482.000'),
('YEARSTARTNYK', '2021-10-19', 'NYK', '1548.000'),
('YEARSTARTOKC', '2021-10-19', 'OKC', '1309.000'),
('YEARSTARTORL', '2021-10-19', 'ORL', '1330.000'),
('YEARSTARTPHI', '2021-10-19', 'PHI', '1609.000'),
('YEARSTARTPHX', '2021-10-19', 'PHX', '1650.000'),
('YEARSTARTPOR', '2021-10-19', 'POR', '1581.000'),
('YEARSTARTSAC', '2021-10-19', 'SAC', '1454.000'),
('YEARSTARTSAS', '2021-10-19', 'SAS', '1478.000'),
('YEARSTARTTOR', '2021-10-19', 'TOR', '1449.000'),
('YEARSTARTUTA', '2021-10-19', 'UTA', '1615.000'),
('YEARSTARTWAS', '2021-10-19', 'WAS', '1495.000');

UNLOCK TABLES;


```{r}
# Goal: Calculate Elo rating for the current date
# 1. Pull all the ELO rating rows from previous day
# 2. Pull NBA game results from previous day
# 3. Calculate the ELO rating for each team (30) for the current date (morning of)
# 4. If team doesn't play, then obviously, just copy forward the value
# 5. Write a new row for each team for the date

# 1. Pull all the ELO rating rows from previous day
localuserpassword <- "BostonCeltics17"
# The connection method below uses a password stored in a variable.
# To use this set localuserpassword="The password of newspaper_search_results_user"

nbaDb <- dbConnect(MariaDB(), user='nba_game_statistics_user', password=localuserpassword, dbname='nba_game_statistics', host='localhost')


# You can fetch all results:
res <- dbSendQuery(nbaDb, "SELECT * FROM elo_rating WHERE DATE = '2021-10-19'")
elo_ratings_yday <- dbFetch(res)
dbClearResult(res)

res <- dbSendQuery(nbaDb, "SELECT * FROM schedule WHERE DATE = '2021-10-19'")
games_yday <- dbFetch(res)
dbClearResult(res)

#dbWriteTable(nbaDb, "schedule", games_df, append=TRUE ) 

# Disconnect to clean up the connection to the database.

dbDisconnect(nbaDb)

print(elo_ratings_yday)
print(games_yday)
```

```{r}
elo_ratings_today <- elo_ratings_yday
elo_ratings_today$DATE <- '2021-10-20'
#print(elo_ratings_today)
#games_yday[1,]$HOME_TEAM

print(games_yday)
print(elo_ratings_today)
for (i in 1:nrow(games_yday)) {
  # calculate_elo_ratings(home_elo, away_elo, home_score, away_score)
  home_elo_y_row <- elo_ratings_yday %>% filter(TEAM == games_yday[i,]$HOME_TEAM)
  away_elo_y_row <- elo_ratings_yday %>% filter(TEAM == games_yday[i,]$AWAY_TEAM)
  
  home_elo_y <- home_elo_y_row$ELO_RATING
  away_elo_y <- away_elo_y_row$ELO_RATING
  
  home_score_y <- games_yday[i,]$HOME_TEAM_SCORE
  away_score_y <- games_yday[i,]$AWAY_TEAM_SCORE
  
  # calculate_elo_ratings <- function(home_elo, away_elo, home_score, away_score)
  elo_results <- calculate_elo_ratings(home_elo_y, away_elo_y, home_score_y, away_score_y)
  elo_ratings_today$ELO_RATING[elo_ratings_today$TEAM == games_yday[i,]$HOME_TEAM] <- elo_results[[1]]
  elo_ratings_today$ELO_RATING[elo_ratings_today$TEAM == games_yday[i,]$AWAY_TEAM] <- elo_results[[2]]

  #print(elo_results)
}


elo_ratings_today <- subset(elo_ratings_today, select = -c(GAME_ID))
print(elo_ratings_today)


```


```{r}
localuserpassword <- "BostonCeltics17"
# The connection method below uses a password stored in a variable.
# To use this set localuserpassword="The password of newspaper_search_results_user"

nbaDb <- dbConnect(MariaDB(), user='nba_game_statistics_user', password=localuserpassword, dbname='nba_game_statistics', host='localhost')

dbWriteTable(nbaDb, "elo_rating", elo_ratings_today, append=TRUE ) 

# Disconnect to clean up the connection to the database.

dbDisconnect(nbaDb)
# Above worked, I calculated the Elo for 10/20 based on game results and elo rating
```

```{r}

```


```{r}

```

```{r}

```


```{r}

```

```{r}

```


```{r}

```

```{r}

```


```{r}

```

```{r}

```


```{r}

```

