---
title: "DATA 624 Homework 1"
author: "Philip Tanofsky"
date: "9/12/2021"
output: html_document
---

```{r warning=F, message=F}
# Import required R libraries
library(fpp3)
```

## Exercise 2.1

Use the help function to explore what the series `gafa_stock`, `PBS`, `vic_elec` and `pelt` represent.

**gafa_stock:** Historical stock prices from 2014-2018 for Google, Amazon, Facebook and Apple. All prices are in $USD. Contains data on irregular trading days.

**PBS:** Monthly tsibble with two values, Scripts for total number of scripts and Cost for cost of the scripts in $AUD

**vic_elec:** Half-hourly tsibble with three values, Demand for total electricity demand in MW, Temperature for the temperature of Melbourne (BOM site 086071) and Holiday for the indicator for if that day is a public holiday.

**pelt:** Hudson Bay Company trading records for Snowshoe Hare and Canadian Lynx furs from 1845 to 1935. This data contains trade records for all areas of the company.

```{r warning=F, message=F}
head(gafa_stock)

head(PBS)

head(vic_elec)

head(pelt)
```

### A.

Use autoplot() to plot some of the series in these data sets.

```{r warning=F, message=F}
autoplot(gafa_stock, Close) +
  labs(title = "Closing Stock Price",
       subtitle = "Google, Amazon, Facebook, Apple",
       y = "Price")

pbs_a10 <- PBS %>%
  filter(ATC2 == "A10" & Concession == "Concessional" & Type == "Co-payments") %>%
  mutate(Cost = Cost/1e6) %>%
  select(Month, Cost)

autoplot(pbs_a10, Cost) +
  labs(title = "Total cost of Medicare Australia prescriptions by Month",
       subtitle = "Data defined as ATC2 is A10, Concessional and Co-payments",
       y = "Cost in millions")

autoplot(vic_elec, Demand) +
  labs(title = "Electricity Demand",
       subtitle = "Victoria, Australia",
       y = "Megawatts")

autoplot(pelt, Hare) +
  labs(title = "Trading records of furs from 1845 to 1935",
       subtitle = "Hudson Bay Company",
       y = "Pelts Traded")
```

### B.

#### What is the time interval of each series?

**gafa_stock:** One day, excluding days the stock market is closed (time interval of [!] indicates irregularity)

**PBS:** One month

**vic_elec:** 30 minutes

**pelt:** One year

## Exercise 2.2

Use `filter()` to find what days corresponded to the peak closing price for each of the four stocks in `gafa_stock`.

```{r warning=F, message=F}
result <- gafa_stock %>% 
             group_by(Symbol) %>%
             filter(Close == max(Close)) %>%
             arrange(desc(Close))
result
```


## Exercise 2.3

Download the file `tute1.csv` from the book website, open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

### A.

You can read the data into R with the following script:

```{r warning=F, message=F}
tute1 <- readr::read_csv("tute1.csv")
View(tute1)
```

### B.

Convert the data to time series

```{r warning=F, message=F}
mytimeseries <- tute1 %>%
  mutate(Quarter = yearmonth(Quarter)) %>%
  as_tsibble(index = Quarter)

head(mytimeseries)
```

### C.

Construct time series plots of each of the three series

```{r warning=F, message=F}
mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y")
```

Check what happens when you don’t include facet_grid().

As seen below, without `facet_grid()`, the three time series are all graphed on the same graph instead of 3 individual graphs. I would argue this provides a more contextualized graph of the relationship between the three separate time series. With a shared y-axis, the relationships between the three time series are more evident than the separated graphs.


```{r warning=F, message=F}
mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line()
```

## Exercise 2.4

The `USgas` package contains data on the demand for natural gas in the US.

### A. 

Install the `USgas` package.

```{r warning=F, message=F}
library(USgas)
```

### B.

Create a tsibble from `us_total` with year as the index and state as the key.

```{r warning=F, message=F}
ts <- us_total

ts <- ts %>%
  as_tsibble(index = year, key = state)

head(ts)
```

### C.

Plot the annual natural gas consumption by state for the New England area (comprising the states of Maine, Vermont, New Hampshire, Massachusetts, Connecticut and Rhode Island).

```{r warning=F, message=F}
ne_ts <- ts %>%
  filter(state == 'Maine' |
           state == 'Vermont' |
           state == 'New Hampshire' |
           state == 'Massachusetts' |
           state == 'Connecticut' |
           state == 'Rhode Island') %>%
  mutate(y = y/1e3)


head(ne_ts)

autoplot(ne_ts, y) +
  labs(title = "Annual natural gas consumption by state",
       subtitle = "New England area",
       y = "Consumption in thousands")
```

## Exercise 2.5

### A.

Download `tourism.xlsx` from the book website and read it into R using `readxl::read_excel()`.

```{r warning=F, message=F}
tourism_data <- readxl::read_excel("tourism.xlsx")

head(tourism_data)
```

### B.

Create a tsibble which is identical to the `tourism` tsibble from the tsibble package.

First, display `tourism` as sample.

```{r warning=F, message=F}
# Output tourism tsibble
head(tourism)
```

Initial thoughts, five columns are the same. Steps: 1: Convert column Quarter into the index. 2: Define Region, State and Purpose as keys.

```{r warning=F, message=F}
# Convert tourism_data to tsibble
tourism_ts <- tourism_data %>%
  mutate(Quarter = yearquarter(Quarter)) %>%
  as_tsibble(index = Quarter, key = c(Region, State, Purpose))

head(tourism_ts)
```

### C.

Find what combination of `Region` and `Purpose` had the maximum number of overnight trips on average.

```{r warning=F, message=F}
tourism_by_reg_pur_ts <- tourism_data %>%
  group_by(Region, Purpose) %>%
  summarise(Trip_Avg = mean(Trips)) %>%
  filter(Trip_Avg == max(Trip_Avg)) %>%
  arrange(desc(Trip_Avg))

head(tourism_by_reg_pur_ts)
```

**Answer:** Combination of region Sydney and purpose Visiting has the maximum number of overnight trips on average per quarter surveyed with a count of 747.

### D.

Create a new tsibble which combines the Purposes and Regions, and just has total trips by State.

```{r warning=F, message=F}
trips_by_state <- tourism_ts %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips)) %>%
  mutate(Quarter = yearquarter(Quarter)) %>%
  as_tsibble(index = Quarter, key = State)

head(trips_by_state)
```

## Exercise 2.8

Monthly Australian retail data is provided in `aus_retail`. Select one of the time series as follows (but choose your own seed value):

```{r warning=F, message=F}
#set.seed(12345678)
set.seed(8675309)

myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))

head(myseries)

dim(myseries)
```

Explore your chosen retail time series using the following functions:

`autoplot()`, `gg_season()`, `gg_subseries()`, `gg_lag()`, `ACF() %>% autoplot()`

```{r warning=F, message=F}
autoplot(myseries, Turnover) +
  labs(title = "Turnover in Queensland Takeaway food services",
       subtitle = "Series ID: A3349767W",
       y = "Turnover")
```

`autoplot()` indicates a increasing trend.

```{r warning=F, message=F}
gg_season(myseries, Turnover) +
  labs(title = "Turnover in Queensland Takeaway food services",
       subtitle = "Series ID: A3349767W",
       y = "Turnover")
```

`gg_season()` indicates a seasonal pattern particularly in the more recent years.

```{r warning=F, message=F}
gg_subseries(myseries, Turnover) +
  labs(title = "Turnover in Queensland Takeaway food services",
       subtitle = "Series ID: A3349767W",
       y = "Turnover")
```

`gg_subseries()` reiterates the finding from `gg_season()`. The blue average line represents the seasonal, yearly nature of the turnover.

```{r warning=F, message=F}
gg_lag(myseries, geom ="point") +
  labs(x = "lag(Turnover, k)")
```

I'll admit the `gg_lag()` is not telling me much besides a strongly positive relationship. My impression is that the graph is speaking more to the overall increase of turnover.

```{r warning=F, message=F}
myseries %>% ACF(Turnover) %>% autoplot()
```

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

I would say an increasing trend is present in the data and the data does follow a seasonal pattern over the course of a given year. The `ACF()` graph almost captures the scallop graphical pattern as identified in the textbook, but not quite. Given the step-down pattern present after lag 12, 24, I do define this data as seasonal on a yearly basis. I would not say a cyclic nature exists for the data.

In wanting to say Yes to a cyclic nature, perhaps an argument could be made for a general 7-year stock-market cycle starting in 1998 through 2014, but I think that's a stretch. I'm not an expert on the Australian economy, but if the Australian economy follows the US economy based on global impacts, then I could believe the dot com bust of late 90s and the 2008 subprime mortgage market crash are being reflected here. But I wouldn't say that's a cycle. I wouldn't say this data captures a true stock-market cycle given the data provided.