---
title: "DATA 624 Homework 1"
author: "Philip Tanofsky"
date: "9/6/2021"
output: html_document
---

```{r warning=F, message=F}
# Import required R libraries
library(fpp3)
```

## Exercise 2.1

Use the help function to explore what the series `gafa_stock`, `PBS`, `vic_elec` and `pelt` represent.

gafa_stock: historical stock prices from 2014-2018 for Google, Amazon, Facebook and Apple

```{r warning=F, message=F}
head(gafa_stock)

head(PBS)

head(vic_elec)

head(pelt)
```

### A.

Use autoplot() to plot some of the series in these data sets.

### B.

What is the time interval of each series?

## Exercise 2.2

Use `filter()` to find what days corresponded to the peak closing price for each of the four stocks in `gafa_stock`.

## Exercise 2.3

Download the file `tute1.csv` from the book website, open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

### A. 

You can read the data into R with the following script:

```{r warning=F, message=F}
tute1 <- readr::read_csv("tute1.csv")
View(tute1)
```

### B. Convert the data to time series

```{r warning=F, message=F}
mytimeseries <- tute1 %>%
  mutate(Quarter = yearmonth(Quarter)) %>%
  as_tsibble(index = Quarter)
```

### C. Construct time series plots of each of the three series

```{r warning=F, message=F}
mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y")
```

Check what happens when you don’t include facet_grid().

## Exercise 2.4

The `USgas` package contains data on the demand for natural gas in the US.

### A. 

Install the `USgas` package.

```{r warning=F, message=F}
library(USgas)
```

### B.

Create a tsibble from `us_total` with year as the index and state as the key.

```{r warning=F, message=F}
ts <- us_total

head(ts)
```

### C.

Plot the annual natural gas consumption by state for the New England area (comprising the states of Maine, Vermont, New Hampshire, Massachusetts, Connecticut and Rhode Island).

## Exercise 2.5

### A.

Download `tourism.xlsx` from the book website and read it into R using `readxl::read_excel()`.

```{r warning=F, message=F}
tourism_data <- readxl::read_excel("tourism.xlsx")

head(tourism_data)
```

### B.

Create a tsibble which is identical to the `tourism` tsibble from the tsibble package.

### C.

Find what combination of `Region` and `Purpose` had the maximum number of overnight trips on average.

### D.

Create a new tsibble which combines the Purposes and Regions, and just has total trips by State.

## Exercise 2.8

Monthly Australian retail data is provided in `aus_retail`. Select one of the time series as follows (but choose your own seed value):

```{r warning=F, message=F}
#set.seed(12345678)
set.seed(8675309)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))
```

Explore your chosen retail time series using the following functions:

`autoplot()`, `gg_season()`, `gg_subseries()`, `gg_lag()`, `ACF() %>% autoplot()`

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?







