---
title: "Technical Report: Determinants of PH in Beverage Process"
subtitle: ""
author: "Alexander Ng, Philip Tanofsky"
date: "Due 12/13/2021"
output:
  tufte::tufte_html: 
    toc: true
    toc_depth: '2'
    highlight: pygments
    number_sections: no
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    df_print: paged
    highlight: tango
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 0
    toc_float: no
    fontsize: 12
editor_options:
  chunk_output_type: inline
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Overview

```{r warning=FALSE, message=FALSE }
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages( library(tidyverse))
suppressPackageStartupMessages(library(kableExtra))
library(cowplot)
library(skimr)

library(RColorBrewer)
library(caret)
library(Cubist)
library(rpart)
suppressPackageStartupMessages(library(gbm))
library(party)
library(tictoc)
tic()
```

```{css, echo=FALSE}
.code-bg {
  background-color: lightgray;
}
```

# Introduction

This technical report tunes multiple models for prediction of beverage PH and identifies determinants of PH among predictors in the manufacturing process.
We train and test 5 models for this purpose including the Cubist, Gradient Boosted Tree, CART, MARS and linear regression models.

The training and test performance were XXXPSTXXX.

# Importing Data

We load the pre-processed datasets for training and testing in this section.
We load the csv file for the training data set below and display the first few column characteristics from the file.

```{r data-load, echo = TRUE}

sdata_train_pp = read_csv("sdata_train_pp.csv", 
                          col_types = cols( .default = "?", id = "i" , BrandCode = "f")) %>% 
  as_tibble() %>% remove_rownames()

sdata_test_pp  = read_csv("sdata_test_pp.csv", col_types = cols( .default = "?", id = "i" , BrandCode = "f")) %>% 
  as_tibble() %>% remove_rownames()

sdata_trainX_pp  = sdata_train_pp %>% select( -PH, -id )
sdata_testX_pp   = sdata_test_pp %>% select(-PH , -id)
```

```{r echo = TRUE}
tuningFull = TRUE   # Set to TRUE if running the lengthy tuning process for all models
```

# Gradient Boosted Trees

We consider the same dataset and variable importance problem using gradient boosted trees.
The `gbm` library and variable importance measures are run within caret training procedure. The gradient boosting approach relies on a weak learner in an additive manner that minimizes the given loss function. 

```{r gbm-tuning, echo = TRUE , warning = FALSE}
set.seed(1027)

gbmControlFull <- trainControl(method = "repeatedcv", 
                           number = 10,  # for debug mode, set this to a low number (1)
                           repeats = 10 ,
                           selectionFunction = "best", 
                           verboseIter = FALSE)

gbmControlLight <- trainControl(method = "repeatedcv", 
                           number = 5,  # for debug mode, set this to a low number (1)
                           repeats = 1 ,
                           selectionFunction = "best", 
                           verboseIter = FALSE)

if(tuningFull == TRUE)
{
   gbmControl = gbmControlFull
   gbmTune = expand.grid( n.trees = seq(250, 1000, by = 250),
                                                  shrinkage = c( 0.02, 0.05,  0.1) ,
                                                  interaction.depth = c(  3, 7) ,
                                                  n.minobsinnode = 15
                                                  )
} else  {
   gbmControl = gbmControlLight
   gbmTune =  expand.grid( n.trees = c( 2000), 
                                                  shrinkage = c( 0.02, 0.05 , 0.1 ) ,
                                                  interaction.depth = c(  3, 7) ,
                                                  n.minobsinnode = 4
                                                  )
}

(gbmTune = caret::train( x = sdata_trainX_pp, 
                          y = sdata_train_pp$PH , 
                          method = "gbm",
                          tuneGrid =  gbmTune ,
                          verbose = FALSE,
                          metric = "RMSE" ,
                          trControl = gbmControl ) )
```

As the model output indicates the greater tree depth outperforms the lower counts and the lower shrinkage value produces a lower (better) RMSE value. The shrinkage, or learning rate, works better with smaller values, yet require longer processing time.

```{r}
gbm_plot <- ggplot(gbmTune) + labs(title = "Gradient Boosted Trees - Tuning")
gbm_plot
ggsave("GradientBoostedTreesTurning.jpg", gbm_plot)
```

The plot of the tuned model clearly shows the better performance with the increased max tree depth but also indicates the lower shrinkage value doesn't guarantee a lower RMSE. For the max tree depth of 3, the lowest shrinkage value does not produce the lowest RMSE for that tree depth.

```{r}
gbm_vi_plot <- ggplot(varImp(gbmTune, numTrees = gbmTune$finalModel$n.trees) ) + labs(title = "Gradient Boosted Trees - Variance Importance")
gbm_vi_plot
ggsave("GbmVarImp.jpg", gbm_vi_plot)
```

The variable importance plot shows the 100% importance of the `MnfFlow` followed by `BrandCode` as the only other variable with an importance above 50%. `Usagecont`, `OxygenFiller`, and `Temperature` round out the top five variables based on importance with scores greater than 25%.

Overall, the final Gradient Boosted model results in an RMSE of approximately 0.13 and an $R^2$ less than 0.5.

# Cubist Model

Now we train a Cubist tree using the `best` selectionFunction on RMSE. The `Cubist` library and variable importance measures are run within caret training procedure.\
The Cubist model is a rules-based approach in which the terminal leaves contains linear regression models. The models are based on predictions defined in the previous nodes of the tree.

```{r cubist-tuning, echo = TRUE, warning=FALSE, error = FALSE, message=FALSE}
set.seed(1027)

cubistControlFull <- trainControl(method = "cv" ,  selectionFunction = "best")
tuneGridFull  <- expand.grid( committees = c( 10, 50, 100 ) ,
                              neighbors = c( 0, 1, 5, 9 )
                                                  ) 

cubistControlLight <- trainControl(method = "cv" ,  selectionFunction = "best")
tuneGridLight <- expand.grid( committees = c( 10, 20 ) , 
                              neighbors = c( 0, 5 )  )

if(tuningFull == TRUE)
{
   cubistControl = cubistControlFull
   cubistGrid = tuneGridFull
} else  {
   cubistControl = cubistControlLight
   cubistGrid = tuneGridLight
}

(cubistTune = caret::train( x = sdata_trainX_pp, 
                          y = sdata_train_pp$PH , 
                          method = "cubist",
                          tuneGrid = cubistGrid ,
                          verbose = FALSE,
                          metric = "RMSE" ,
                          trControl = cubistControl ) )
```

The Cubist model results provide a good RMSE value and $R^2$ value with a higher number of committees and neighbors.

The line plot confirms the better model per RMSE with more neighbors and increasing number of committees.

```{r}
cubist_plot <- ggplot(cubistTune) + labs(title="Cubist Model Tuning")
cubist_plot
ggsave("CubistTuning.jpg", cubist_plot)
```

The variable importance plot identifies `MnfFlow` as the most valuable variable followed by `BallingLvl`, `Balling`, `AlchRel`, and `PressureVacuum` to round out the top 5 variables. Overall, 8 variables resulted in over 50% importance indicating those variables are contained in over 50% of the rules defining the Cubist model.

```{r}
cubist_vi_plot <- ggplot(varImp(cubistTune), top = 20) + labs(title = "Cubist Model Variable Importance")
cubist_vi_plot
ggsave("CubistVarImp.jpg", cubist_vi_plot)
```


```{r eval=F}
cubistTune$finalModel
```

The final Cubist model confirms the high usage rate of the variable `MnfFlow`.

```{r}
cubistTune$finalModel$usage %>% arrange( desc(Model+ Conditions)) %>% 
  kable(caption = "Cubist Predictor Usage") %>% kable_styling(bootstrap_options = c("hover", "striped"), position = "left")
```

The `splits` plot indicates the usage of each variable by committee/rule. The values are normalized and highlight the high usage rate of `MnfFlow` along with the use of greater than or less than inequalities for `MnfFlow`.

```{r}
cubist_splitsplot <- dotplot( cubistTune$finalModel, what = "splits" )
cubist_splitsplot
```

```{r}
# Output to file
jpeg(file="CubistSplitsPlot.jpg")
cubist_splitsplot
dev.off()
```

The `dotplot` allows us to see the distribution of the coefficient parameters for each linear model associated with a committee/rule.\
When a predictor has more dots, it is used in more models which suggests it is influential. The color density is proportional to the variable importance.
The sign of the coefficient tells us whether the marginal effect of the predictor on the response `PH`. However, since Cubist uses many models, the concept of the `sign` of a predictor coefficient is unclear.\
The x-axis value of the dot indicates the true coefficient in the original units unlike the above plot which normalizes the x-axis.

```{r}
cubist_dotplot <- dotplot( cubistTune$finalModel, 
         what = "coefs", 
         between = list(x = 0.2, y = 0.5) , 
         scales = list(x = list(relation = "free"),  
                       y = list(cex = 0.25)  ) )
cubist_dotplot
jpeg(file="CubistDotPlot.jpg")
cubist_splitsplot
dev.off()
```

The boxplot below which we devised constructs a distribution of coefficients for each predictor.
We interpret the sign of the predictor $X$ marginal effect to be the sign of the median of the distribution of coefficients pf $X$ over those models which include $X$.
The predictors have been sorted by their median.

The plot provides clarity regarding the impact of a given variable as the median is a good estimate of the sign of the variable coefficient. The `AlchRel` variable box indicates a positive relationship with the dependent variable `PH`. The most important variable `MnfFlow` box results in a negative coefficient, and thus an inverse effect on `PH`. For all the rules in the final Cubist model, the understanding of the coefficient values for each variable provides a clear relationship to the dependent variable despite the obtuse nature of the Cubist model construction.

Top 10 importance:
MnfFlow: Negative
BallingLvl: Positive
Balling: Negative
AlchRel: Positive
PressureVacuum: N
Density: N
OxygenFilter: N
AirPressurer: N
BowlSetpoint: P
HydPressure3: P

LM, coeff pos or neg:
MnfFlow: N-check
BallingLvl: P-check
Balling: N-check
AlchRel: P-check
PressureVacuum: N-check
Density: N-check
OxygenFilter: N-check
AirPressure: N-check
BowlSetpoint: P-check
HydPressure3: P-check

```{r}
cubistTune$finalModel$coefficients %>% 
  rownames_to_column(var="id") %>% 
  select(-committee, -rule) %>% 
  pivot_longer(!id, names_to = "predictor", values_to = "coef" ) %>% 
  filter( !is.na(coef)) %>% 
  filter( predictor != '(Intercept)' )  %>%
  filter(abs(coef) < 20 ) -> coef_piv

cubist_boxplot <- coef_piv %>% 
  ggplot() + 
  geom_boxplot(
    aes(x=reorder(predictor, coef, FUN = median, na.rm = TRUE ),  # order the boxplots by median
        y = coef ), outlier.shape = NA ) +  # strip out the outliers
  coord_flip(ylim=c(-.5,.5)) +
  labs(title = "Distribution of Coefficients of Predictors", 
       subtitle = "from Cubist Committee/Rules",
       x = "Predictors" ,
       y = "Coefficient in Linear Model"
       )

cubist_boxplot
ggsave("CubistBoxPlot.jpg", cubist_boxplot)
```

Overall, the final Cubist model results in an RMSE under 0.1 and an $R^2$ just short of 0.7.

# CART Tree

Using the `rpart` library and its training methods, we consider the best CART based on the `RMSE` metric. The CART approach follows a classification and regression tree methodology.

```{r cart-tuning, echo = TRUE, warning = FALSE}
set.seed(10393)

library(rpart)
cartGrid = expand.grid( maxdepth = seq( 1, 20, by = 1 )  )

cartControl <- trainControl(method = "boot", 
                            number = 30,  
                            selectionFunction = "best")

(rpartTune = train( x = sdata_trainX_pp, 
                    y = sdata_train_pp$PH ,
                   method = "rpart2" ,
                   metric = "RMSE" ,
                   tuneGrid = cartGrid ,
                   trControl = cartControl ) )
```

The model output shows increasing tree depth leads to improved RMSE and $R^2$ values.

The below plot displays the improving RMSE value with the increasing tree depth with an optimal tree depth of 19.

```{r}
cart_tuning_plot <- ggplot(rpartTune) + labs(title="CART Tree Tuning")
cart_tuning_plot
ggsave("CartTuningPlot.jpg", cart_tuning_plot)
```

Once again, the variable `MnfFlow` is the most important variable with an importance rating of 100%. Six variables achieve an importance score of greater than 50% with `BrandCode`, `Temperature`, and `CarbRel` each achieving greater than 75% importance.

```{r}
cart_vi_plot <- ggplot(varImp( rpartTune), top = 20 ) + labs(title = "CART Tree Variable Importance")
cart_vi_plot
ggsave("CartVarImp.jpg", cart_vi_plot)
```

The CART tree plot of the final model reaffirms the importance of the variable `MnfFlow` as the initial node splits on `MnfFlow`. One of the second level nodes is `BrandCode`, another highly important variable to the CART model. Eight ff the top 10 most important variables in the Cubist model are found in the top 15 variables of the CART model. This overlap of variable importance shows a consistency across the different model approaches.


CART top 20:
MnfFlow :IN CUBIST TOP 10
BrandCode
Temperature
CarbRel
BallingLvl :IN CUBIST TOP 10
FillerSpeed
HyPressure3 :IN CUBIST TOP 10
Usagecont
BowlSetpoint  :IN CUBIST TOP 10
FillerLevel
OxygenFiller :IN CUBIST TOP 10
AirPressurer :IN CUBIST TOP 10
PressureVacuum :IN CUBIST TOP 10
Balling :IN CUBIST TOP 10
FillPressure
HydPressure1
PSCFill
CarbPressure1
HydPressure2
PressureSetpoint



```{r}
library(rpart.plot)
library(partykit)

rpart.plot(rpartTune$finalModel, digits = 3)
```

Overall, the final CART model results in an RMSE of approximately 0.13 and an $R^2$ less than 0.5.

# MARS Model

Now we train a MARS model based on the implementation of earth from the `caret` library. The MARS approach is a non-linear regression model that doesn't rely on the original variables but instead on surrogate features.

The tuned MARS model defines the derived features as linear combinations of the initial variables.

```{r  mars-tuning , message = FALSE, warning= FALSE, error= FALSE, class.source = 'code-bg'}

if( tuningFull == TRUE )
{
    marsGrid = expand.grid( .degree = 1:2, .nprune = seq(2,27, by=5) )
    marsControl = trainControl(method = "cv", 
                     number = 10, 
                
                     selectionFunction = "best", 
                     verboseIter = FALSE)
} else {
    marsGrid = expand.grid( .degree = 1:2, .nprune = seq(2,25, by = 10 ) )
    
    marsControl = trainControl(method = "cv", 
                     number = 10, 
                     
                     selectionFunction = "best", 
                     verboseIter = FALSE)
}
    

set.seed(1000)

marsTuned = caret::train(x = sdata_trainX_pp , 
                         y = sdata_train_pp$PH, 
                         method = "earth" ,
                         metric = "RMSE" , 
                         tuneGrid = marsGrid ,
                         trControl = marsControl )
```


```{r}
marsTuned$finalModel

summary(marsTuned)
```

The plot of the final MARS model indicates increasing the number of terms leads to a lower RMSE.

```{r}
mars_tuning_plot <- ggplot(marsTuned) + labs(title="MARS Model Tuning" )
mars_tuning_plot
ggsave("MarsTuningPlot.jpg", mars_tuning_plot)
```

Once again, the variable `MnfFlow` achieves an importance score of 100%. Only 3 variables reach the score of 50%, the aforementioned `MnfFlow`, `BrandCodeC`, and `AlchRel`.\
The variable `MnfFlow` is influential because the variable effects many other variables in the final model. Due to the numerous linear combinations including `MnfFlow`, the relationship to `PH` is difficult to ascertain. The `MnfFlow` variable requires more investigation into the variable interactions to determine the impact to `PH`.

```{r}
mars_vi_plot <- ggplot( varImp(marsTuned), 20 ) + labs(title = "MARS Model Variable Importance")
mars_vi_plot
ggsave("MarsVarImp.jpg", mars_vi_plot)
```

Overall, the MARS model reaffirms the importance of the variable `MnfFlow` and also results in an RMSE and $R^2$ in line with the previous models.

# Linear Regression

We also run a simple OLS regression on PH in terms of the scaled and centered variables.
While its predictive performance is weaker than the other models on the training set, the coefficients are much easier to interpret.

```{r}
set.seed(123)

train.control <- trainControl( method = "cv", number = 5)

linearTune = train( PH ~ . -id, 
                    data = sdata_train_pp, 
                    method = "lm", 
                    tuneGrid = data.frame(intercept = TRUE),
                    trControl = train.control)

summary(linearTune$finalModel)
```

The most significant variables are `MnfFlow`, `CarbPressure`, `HydPressure3`, `Temperature`, `Usagecont`, `Balling`, `PressureVacuum`, `OxygenFiller`, `BowlSetpoint`, `BallingLvl`, and `BrandCode`. Seven of these 11 variables are found in the top 10 most important variables of the Cubist model. Upon closer evaluation, the median value of the variable coefficients from the top 10 of the Cubist model match the sign of the coefficient value in the linear regression model. This finding further supports the assessment of the Cubist model variable impacts on the dependent variable `PH`.

We summarize the plot of coefficients below.

```{r}
par(mfrow=c(2,2))

lm_final_model_plot <- plot(linearTune$finalModel)
lm_final_model_plot
ggsave("LinearModelFinalPlot.jpg", lm_final_model_plot)
```

```{r}
hist(linearTune$finalModel$residuals)
```

The diagnostic plots of the OLS model suggest that the histogram of residuals is relatively symmetric and looks normal.\
The QQ plot suggests the model fit does not fully capture the fat tails possibly due to a small number of outliers.

Overall, the linear regression model does not perform as well as the Cubist model but does provide a valuable sanity check and baseline to the results from the more advanced models.

# Model Selection

Based on the above results, we summarize their training and test RMSE and $R^2$ below.

The joint results are tabulated below.

```{r compare-model-results , echo = TRUE }

# Make a list of model results collect them from caret for both training and test data and compare them jointly in a kable()

model_list = list( rpartTune , marsTuned , gbmTune, cubistTune, linearTune)

model_stats = data.frame()

for( modelObj in model_list )
{
    if( modelObj$method != 'lm' ){
        pred_data <- as.numeric( predict( modelObj, newdata=sdata_testX_pp ))[]
    }
    else
    {
        pred_data <- as.numeric( predict( modelObj, newdata=sdata_test_pp ))
    }
    output <- data.frame( modelName = modelObj$method ,
                     trainRMSE = getTrainPerf(modelObj)[1,1] ,
                     trainR2   = getTrainPerf(modelObj)[1,2] ,
                     testRMSE  = caret::RMSE( pred_data,  sdata_test_pp$PH) ,
                     testR2    = caret::R2(   pred_data,  sdata_test_pp$PH) 
                     )

    model_stats <- rbind(model_stats, output)
}

```

```{r}
results_table <- model_stats %>% as_tibble() %>% 
  arrange( testRMSE) %>%
  kable( digits = 3, caption = "Model Results Comparison") %>%
  kable_styling(bootstrap_options = c("hover", "striped"), position = "left")

results_table

save_kable(results_table, "ModelResultsTable.jpg")
```

We may also inspect the plot of the predicted versus the observed data for the 5 models.
It is clear below that cubist outperforms all others on both the training set in the upper row and also the test data on the lower rows.

```{r}
predVals = extractPrediction(list(marsTuned, rpartTune, cubistTune, gbmTune), testX = sdata_testX_pp , testY = sdata_test_pp$PH )

# Attempting to hack into the extract prediction object with the LM data
# Cols: obs pred model dataType object
pred_lm =  predict(linearTune, newdata = sdata_test_pp)
pred_lm_train <- linearTune$finalModel$fitted.values

# Create placeholder data for the training values
model_l <- rep("LM", length = 2055)
dt_l <- rep("Training", length = 2055)
obj_l <- rep("Object9", length = 2055)

# Create dataframe of the training data of LM
lm_train_preds <- as.data.frame(cbind(obs=sdata_train_pp$PH, pred=pred_lm_train, model=model_l, dataType=dt_l, object=obj_l), row.names = NULL)

# Create placeholder data for the test values
model_l <- rep("LM", length = 512)
dt_l <- rep("Test", length = 512)
obj_l <- rep("Object10", length = 512)

# Create dataframe of the test data of LM
lm_test_preds <- as.data.frame(cbind(obs=sdata_test_pp$PH, pred=pred_lm, model=model_l, dataType=dt_l, object=obj_l), row.names = NULL)

# Combine all data for plotting
all_preds <- rbind(predVals, lm_train_preds, lm_test_preds)

# Ensure obs and preds are numerics
all_preds$obs <- as.numeric(all_preds$obs)
all_preds$pred <- as.numeric(all_preds$pred)

# Plot using Caret built-in reporting.
# --------------------------------------
obsVsPredsPlot <- plotObsVsPred(all_preds)
obsVsPredsPlot
#ggsave("ObsVsPredsPlot.jpg", obsVsPredsPlot)
```


# Discussion

We conclude that the best model is Cubist but the multilinear regression model, MARS and CART models provide useful insight on the main predictors.

-   We would recommend the use of the Cubist model for prediction
-   However, the OLS (ordinary least squares) linear regression help us to understand the sign and magnitude of the most important variables.
-   The MARS model helps to detect non-linearity and predictor interaction in prediction.

The top 5 variables that impact PH are:

-   `MnfFlow` - this appears negatively related to `PH` level based on OLS, MARS and Cubist median plot.

-   `BallingLvl` is positively related to `PH` level.

-   `PreserveVacuum` is negatively related to `PH.`

-   `AlchRel` is positively related to `PH`.

-   `Balling` is negatively related to `PH`.

Lastly, we observe that Brand is probably not a manufacturing predictor that can be controlled.
Mostly likely Brand refers to a recipe for the final product.
However, knowing the desired optimal `PH` for each Brand may help us control the most influential variables that control `PH`, thereby providing better quality control.
While there are over 30 predictors in the dataset, we recommend looking at the top 10 most influential predictors.
Most of the models agree substantially on which predictors to include in the top 10.

For example, the linear regression model coefficients agrees with the sign and variable importance of the top 10 predictors of the Cubist model.
Importance predictors of the OLS model has the highest statistical significance - consistent with variable importance measure of `caret`.

# Final Predictions

```{r warning=F, error=F}
# Combine the initial training and test datasets into 1
sdata_trainfull_pp <- rbind(sdata_train_pp, sdata_test_pp)
sdata_trainfullX_pp <- sdata_trainfull_pp %>% select(-PH , -id)

edata_test_pp = read_csv("edata_test_pp.csv", 
                          col_types = cols( .default = "?", id = "i" , BrandCode = "f")) %>% 
  as_tibble() %>% remove_rownames()

# Train the Cubist model
set.seed(1027)

if(tuningFull == TRUE)
{
   cubistControl = cubistControlFull
   cubistGrid = tuneGridFull
} else  {
   cubistControl = cubistControlLight
   cubistGrid = tuneGridLight
}



(cubistTune = caret::train( x = sdata_trainfullX_pp, 
                          y = sdata_trainfull_pp$PH , 
                          method = "cubist",
                          tuneGrid = cubistGrid ,
                          verbose = FALSE,
                          metric = "RMSE" ,
                          trControl = cubistControl ) )

# Make predictions on the provided dataset
final_pred_data <- as.numeric( predict( cubistTune, newdata=edata_test_pp ))

# Output prediction performance
full_cubist_train_rmse <- getTrainPerf(cubistTune)[1,1]
full_cubist_train_r2   <- getTrainPerf(cubistTune)[1,2]

full_cubist_train_rmse
full_cubist_train_r2

# Output the results to file
# TODO
```

```{r}
final_pred_data
```

# Code

We summarize all the R code used in this project in this appendix for ease of reading.

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```

Tuning full is `r tuningFull`.


9am on 12/8/2021
Tuning full is FALSE. (589.958 sec elapsed)

```{r}
toc()
```